Boosted Regression Trees for Predictive Auto-Tuning

### 2011-12-14 paper/draft1.tex

>>>>

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

% say what we said
In this paper, we've demonstrated a boosted regression tree-based auto-tuning
method, wherein empirical performance data is used to train a machine learning
model of performance for an instrumented GPU kernel.  In contrast to
traditional model-based auto-tuning, where an explicit model of performance is
built on the basis of an understanding of hardware inner workings, and
empirical auto-tuning, where an exhaustive set of implementation configurations
are tried, the present approach generates, from scratch, a model of kernel
performance on the basis of timing data from a user-definable number of kernel
evaluations.  This approach allows significant flexibility to navigate
tradeoffs between offline and run-time costs, and final auto-tuning
performance.  Importantly, this method treats kernels as black-boxes, allowing
the user to auto-tune in the absence of deep knowledge of hardware details
(which may even been unknowable, in the case of hardware not available at the
time of kernel creation).  This approach also frees auto-tuning performance
from a strong dependency on the accuracy (or inaccuracy) of a pre-defined
analytical model.

An important use case for the tools described here is in the development of
user-facing numerical libraries.  Such libraries are a critical component of
scientific computing infrastructure, since they abstract away implementation
details and make algorithms available to a much wider audience.  However, the
abstraction provided by libraries represents a double-edged sword: one hand,
\emph{using} the library is easier, because it presents a unified abstraction
of related functionality.  However, at the same time, any given library routine
might represent a wide range of substantially different problem configurations,
each with distinct computational issues and bottlenecks.  Auto-tuning has long
provided a solution that finesses these two issues, providing multiple
implementations under the hood, for multiple problem settings, and then using
heuristics or explicit, hand-crafted models to select the appropriate
implementation for a given set of inputs.  The development of such auto-tuned
libraries, while extremely successful, is also very difficult.  The
machine-learning-based techniques described here provide a middle ground, where
a library implementor can simply create an instrumented kernel, and allow
generic tools to automatically generate appropriate auto-tuned implementations,
with small (and controllable) run-time costs.

% DDC: the following text looks like random dumping ground

% We're characterizing hardware by a 1-of-N feature vector, simply describing
% which hardware is the current hardware.
% To make better use of auto-tuning data from other platforms, it would be more
% useful to have precise and descriptive features such as: what compute
% capability is present, how many cores are active, what is the bandwidth
% between the various kinds of memory, and how much of various kinds of memory
% is present.  With these features a model-assisted auto-tuning approach
% might be able to make very good guesses on hardware for which no auto-tuning
% has ever been done.
% 
% In a complete implementation of Eq.~\ref{eq:z}, there would be parameters
% related to the physical layout (e.g. strides) of $x$, $f$ and $z$ arrays in
% addition to the mathematical parameters of heights and widths and so on,
% so the total number of filterbank correlation computations that a user might
% be able to demand is astronomical.
% A typical coping mechanism would be to cast an arbitrary problem configuration into a more
% standardized form, such as by copying inputs into aligned memory buffers with
% appropriate padding, and then choosing an appropriate blocking strategy for
% the computation. At that point, auto-tuning efforts can be focussed on the
% kernel for each blocked strategy. However, on GPU hardware, the cost of
% aligning the inputs can be relatively large.  In future work we hope to apply
% our model-based technique to auto-tune in full problem configuration space, so
% that we can optimize as much of the computation as possible within the
% predictive auto-tuning framework (XXX name).
% 
% Platform Space: XXX
% 
% XXX: GCG chapter pg 13 notes that among the grid search, the best parameter
% setting is different for the various platforms.
% 
% XXX: so far we only tried 3 cards... so we're using a 1-of-N feature.
% 
% 
% %\renewcommand{\bibsection}{\section*{Bibliography}}
% %\setlength{\bibsep}{5pt}
% 
% 
% 
% One might expect then, that it would be easy to implement a library providing
% this operation -- this hypothetical library could have quite a simple interface
% following the spirit of FFTW or a BLAS routine:
% \begin{verbatim}
%     gpu_fbcorr(
%         img_shape, filters_shape,
%         img, filters, output)
% \end{verbatim}
% However, as shown in \citet{pinto+cox:2011gcg} and numerous articles on
% related stencil operations XXX, it is challenging to provide an implementation
% or even an implementation strategy that provides satisfactory performance
% across the range of inputs (shapes, physical layouts) that occur even in
% typical usage.
% 
% Current state-of-the art is to do empirical auto-tuning.
% XXX \vspace{12pt}
% 
% Increasingly, machine learning and statistical inference strategies are
% finding application in compiler technology.
% XXX \vspace{12pt}
% 
% machine learning
% 
% This paper shows that a kernel with several  boosted
% 
% 
% There is a natural function from
% problem space x implementation space x platform space to runtime:
% how much wall time elapses on the given platform when solving the given
% problem with the given implementation.
% 
% Auto-tuning is a family of empirical techniques for finding the implementation
% that minimizes that runtime function, when the platform and the problem
% configuration are given.
% 
% XXX: some math, or possibly a picture with the three boxes that was on Pinto's
% whiteboard.
% 
% This paper is about different ways of doing the minimization.
% For a particular problem (filterbank correlation) we
% look at a parameterized kernel implementation (from \citet{pinto+cox:2011gcg})
% and a few hardware platforms, and compare
% random search, grid search, and a hill-climbing strategy with a good, but
% generic, reference implementation.
% Then we show that we can model the LSM function with a regression tree
% well enough to usefully perform the argmin of Equation~\ref{eq:lsm_argmin}
% on the model rather than the actual system. Whereas it may take several
% minutes to perform a hill-climbing or grid search in the original system,
% a hill-climbing search in the model requires a small fraction of a second.
% Model-based autotuning makes it possible to simulate auto-tuning quickly and
% accurately enough to be used within a single library call of the computational
% routine.
% 
% Best arithmetic density
% 
% A bandwidth analysis that takes the blocking structure into account would
% reveal that larger $H$ and $W$ require each block to read a larger input
% region, and store it in a larger shared memory region, which reduces the
% potential for high occupancy.
% 
% Filterbank correlation represents a difficult code-optimization problem
% because of the number of interacting constraints: a large value of $P$ reduces
% arithmetic density, but a small value of $P$ requires that each thread do more
% work, and makes global memory access latency harder to hide. (XXX: is there a
% better reason?)
% 
% 
% The kernel used to perform
% filterbank correlation was parametrized by 10 parameters, some of which
% were binary and others of which were integer-valued. The full configuration
% space included 12000 kernels.
% XXX: how to make sense of these paramters without code listing?
% 
% XXX: What to all the parameters mean?
% 
% XXX: how many configurations are in the configuration space
% 
% XXX: what was the reference kernel, and how was it chosen?
% 
% XXX: analysis of READ traffic, WRITE traffic, DATA re-use and CACHEing
% strategies. FLOPS per write etc.
% 
% XXX: Why not use FFT: convolution in spectral domain?
% 
% XXXX: Talk about memory transfer requirements vs. speed vs. copy...
% 
% we believe that our approach is the most practial one to get high perf on
% recent hardware, since there is no consensus on hw features (and it won't
% change soon)
% 
% XXX: deliver significant performance increases at very low run-time cost (less
% than a second if the input domain has not been seen before, zero if it
% has, since it will hit the cache) compared with the reference
% implementation
% 
% XXX always delivered good performance: optimality without loss of productivity?
% 
% XXX such machine learning derived model-based empirical auto-tuner hold great
% promise for dev productivity, portability, performance on future multi-core hw
% with different architectural features
% 
% XXX frameworks for predicting performance exist \cite{XXX Hwu An Adaptive
% Performance Modeling Tool for GPU Architectures}
% but for non-GPU architectures
% 
% auto-tuning required hardward expertise *and* application-domain knowledge, we
% want to remove these restrictions ultimately we want a ml-based approach to
% code optimization that will where a set of generic but deep code
% transformations, could be templated and injected in any domain-specific
% library. this will reduce the need for specialized expertise across the range
% of targeted computing systems.
